---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

nformat <- function(x)
  format(x, big.mark = ',')
```


# mcgibbsit

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/mcgibbsit)](https://CRAN.R-project.org/package=mcgibbsit)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

Implementation of Warnes & Raftery's MCGibbsit run-length diagnostic
for a set of (not-necessarily independent) MCMC samplers.  

It combines the estimate error-bounding approach of the Raftery and Lewis MCMC
run length diagnostic (`gibbsit`) with the between verses within chain approach of
the Gelman and Rubin MCMC convergence diagnostic.

## Installation

Install the most recent release from `CRAN`:
``` r
install.packages("mcgibbsit")
```

Or the current development version from `github`:
```r
if(!require("remotes"))
  install.packages("remotes")
  
remotes::install_github('r-gregmisc/mcgibbsit')
```

## Example

This basic example constructs a dummy set of files from an *imaginary* MCMC sampler and shows the results of running `mcgibbsit` with the default settings.

```{r}
library(mcgibbsit)

###
# Create example data files for 20 independent chains
###

set.seed(42)
tmpdir <- tempdir()

nsamples <- 200

# Define a function to generate the output of our imaginary MCMC sampler
gen_samples <- function(run_id)
{
  x <- matrix(nrow = nsamples+1, ncol=4)
  colnames(x) <- c("alpha","beta","gamma", "nu")
  
  x[,"alpha"] <- exp(rnorm (nsamples+1, mean=0.025, sd=0.025))
  x[,"beta"]  <- rnorm (nsamples+1, mean=53,    sd=14)
  x[,"gamma"] <- rbinom(nsamples+1, 20,         p=0.15) + 1
  x[,"nu"]    <- rnorm (nsamples+1, mean=x[,"alpha"] * x[,"beta"], sd=1/x[,"gamma"])
#'
  # induce serial correlation of 0.25
  x <- 0.75 * x[2:(nsamples+1),] + 0.25 * x[1:nsamples,]

  # induce ~50% acceptance rate
  accept <- runif(nsamples) > 0.50
  for(i in 2:nsamples)
    if(!accept[i]) x[i,] <- x[i-1,]

  write.table(
    x,
    file = file.path(
      tmpdir,
      paste("mcmc", run_id, "csv", sep=".")
      ),
    sep = ",",
    row.names = FALSE
  )
}
```

First, we'll generate and load only a 3 runs of length 200:

```{r}
# Generate and load 3 runs 
for(i in 1:3)
  gen_samples(i)
  
mcmc.3 <- read.mcmc(
  3, 
  file.path(tmpdir, "mcmc.#.csv"), 
  sep=",",
  col.names=c("alpha","beta","gamma", "nu")
  )
```

```{r trace_density_plot_3,fig.width=8,fig.height=8}
# Trace and Density Plots
plot(mcmc.3)
```

Now run `mcgibbsit` to determine the necessary total number of MCMC samples to to provide accurate 95% posterior confidence region estimates for all four of the parameters:
```{r}
# And check the necessary run length 
mcg.3 <- mcgibbsit(mcmc.3)
print(mcg.3)
```

**Interpretation**: `mcgibbsit` estimated that a total of required number of samples is `r nformat(max(mcg.3$resmatrix[,"Total"]))`, which is less than we've generated so far.

Just to be safe we'll add 7 more runs, each of length 200, for a total of  2,000 samples:

```{r}
# Generate and load 7 more runs 
for(i in 3 + (1:7))
  gen_samples(i)
  
mcmc.10 <- read.mcmc(
  10, 
  file.path(tmpdir, "mcmc.#.csv"), 
  sep=",",
  col.names=c("alpha","beta","gamma", "nu")
  )
```

```{r trace_density_plot_10,fig.width=8,fig.height=8}
# Trace and Density Plots
plot(mcmc.10)
```

Now run `mcgibbsit` to determine the necessary number of MCMC samples:
```{r}
# And check the necessary run length 
mcg.10 <- mcgibbsit(mcmc.10)
print(mcg.10)
```

**Interpretation**: `mcgibbsit` now estimates that a total of required number of samples is `r nformat(max(mcg.10$resmatrix[,"Total"]))` (this is slightly fewer than before because the the larger number of samples allowed more accurate estimates of the variances and correlations).  Since we we have already generated `r nformat(mcg.10$nchains * mcg.10$len)` samples, we do not need to perform any additional runs of our MCMC sampler.

We can now calculate the posterior confidence regions for each of the parameters.

```{r}
summary(mcmc.10)
```

